---
title: "Filter"
output: html_document
---

### Set Up

Run for about 100000 schedules at a time unless you have [a lot]{.underline} of RAM.

```{r}
# Load library and some data
library(tidyverse)

# We 
data <- read.csv("CSVNAME.CSV")  

# Make data numeric and add a index for the team number
data <- cbind(index =(seq_len(nrow(data))-1) %% 11, data)
data <- data %>%
  mutate(across(everything(), as.numeric))

```

### Dataframes

Split into different dataframes for easier computation

```{r}
# Opponents
Sched_Track <- data %>%
  select(c(Schedule, index, N1_Opp, C1_Opp, N2_Opp, C2_Opp, N3_Opp, C3_Opp, N4_Opp, C4_Opp, C5_Opp, C6_Opp, N5_Opp, C7_Opp, N6_Opp, C8_Opp, C9_Opp, C10_Opp, N7_Opp, C11_Opp))
  
# Home or away
Home_Track <- data %>%
  select(c(Schedule, index, N1_Home, C1_Home, N2_Home, C2_Home, N3_Home, C3_Home, N4_Home, C4_Home, C5_Home, C6_Home, N5_Home, C7_Home, N6_Home, C8_Home, C9_Home, C10_Home, N7_Home, C11_Home))

# Conference Home or away
Home_Track_Con <- data %>%
  select(c(Schedule, index, C1_Home, C2_Home, C3_Home, C4_Home, C5_Home, C6_Home, C7_Home, C8_Home, C9_Home, C10_Home, C11_Home))

# Non conference home or away
Home_Track_Non <- data %>%
  select(c(Schedule, index , N1_Home, N2_Home, N3_Home,N4_Home,N5_Home,N6_Home, N7_Home))
```

### Physical Distance

Get distance each team travels across the schedule

```{r}
data <- data %>%
group_by(Schedule, index) %>%
  # Noncon distance (*0 if actually at home)
  mutate(NonCon_Dist = sum(N1_Dist * (N1_Home == 0)   + 
                            N2_Dist * (N2_Home == 0)  + 
                            N3_Dist * (N3_Home == 0)  + 
                            N4_Dist * (N4_Home == 0)  + 
                            N5_Dist * (N5_Home == 0)  + 
                            N6_Dist * (N6_Home == 0)  + 
                            N7_Dist * (N7_Home == 0) 
                            , na.rm = TRUE) ,
         # Condistance
         Con_Dist = sum(C1_Dist * (C1_Home == 0)   + 
                           C2_Dist * (C2_Home == 0)+ 
                           C3_Dist * (C3_Home == 0)+ 
                           C4_Dist * (C4_Home == 0)+ 
                           C5_Dist * (C5_Home == 0)+ 
                           C6_Dist * (C6_Home == 0)+ 
                           C7_Dist * (C7_Home == 0)+ 
                           C8_Dist * (C8_Home == 0)+ 
                           C9_Dist * (C9_Home == 0)+ 
                           C10_Dist * (C10_Home == 0)+ 
                           C11_Dist * (C11_Home == 0)
                           , na.rm = TRUE),
         # Total distance
         Total_Dist = NonCon_Dist + Con_Dist)
```

### Bye Week Distance / Opp Distance

FInd distance between bye weeks and each opponent (if there is a non-conference and conference)

```{r}
# Function that returns the number of weeks between repeats (for either bye = -1 or a team number)
bye_gap <- function(row, value) {
  
  indices <- which(trimws(row) == value)
  
  if (length(indices) < 2) {
    return(NA) 
  }
  
  gaps <- diff(indices) - 1
  
  return(paste(gaps, collapse = ", "))
}

# Use function for bye and every team number
Sched_Track$Games_Between_Bye <- apply(Sched_Track[, c(3:20)], 1, bye_gap, value = -1)
Sched_Track$Games_Between_0 <- apply(Sched_Track[, c(3:20)], 1, bye_gap, value = 0)
Sched_Track$Games_Between_1 <- apply(Sched_Track[, c(3:20)], 1, bye_gap, value = 1)
Sched_Track$Games_Between_2 <- apply(Sched_Track[, c(3:20)], 1, bye_gap, value = 2)
Sched_Track$Games_Between_3 <- apply(Sched_Track[, c(3:20)], 1, bye_gap, value = 3)
Sched_Track$Games_Between_4 <- apply(Sched_Track[, c(3:20)], 1, bye_gap, value = 4)
Sched_Track$Games_Between_5 <- apply(Sched_Track[, c(3:20)], 1, bye_gap, value = 5)
Sched_Track$Games_Between_6 <- apply(Sched_Track[, c(3:20)], 1, bye_gap, value = 6)
Sched_Track$Games_Between_7 <- apply(Sched_Track[, c(3:20)], 1, bye_gap, value = 7)
Sched_Track$Games_Between_8 <- apply(Sched_Track[, c(3:20)], 1, bye_gap, value = 8)
Sched_Track$Games_Between_9 <- apply(Sched_Track[, c(3:20)], 1, bye_gap, value = 9)
Sched_Track$Games_Between_10 <- apply(Sched_Track[, c(3:20)], 1, bye_gap, value = 10)
```

### Consecutive

Explore home and away streaks for teams

```{r}
# Function the calculates the maximum number of home or away games a team plays in a row
consec_games <- function(x, value) {
  
  r <- rle(x)
  
  lengths <- r$lengths[r$values == value]

  if (length(lengths) > 0) {
    
    return(max(lengths))
    
  } else {
    return(0)
  }
}

# Apply to conference games, noncon games, and the schedule as a whole
Home_Track$Max_Away_Streak_Total <- apply(Home_Track[, c(3:20)], 1,  
                                          consec_games, value = 0)

Home_Track$Max_Home_Streak_Total <- apply(Home_Track[, c(3:20)], 1, 
                                          consec_games, value = 1)

Home_Track_Con$Max_Away_Streak_Con <- apply(Home_Track_Con[, c(3:13)], 1, 
                                          consec_games, value = 0)

Home_Track_Con$Max_Home_Streak_Con <- apply(Home_Track_Con[, c(3:13)], 1, 
                                          consec_games, value = 1)

Home_Track_Non$Max_Away_Streak_Non <- apply(Home_Track_Non[, c(3:9)], 1, 
                                          consec_games, value = 0)

Home_Track_Non$Max_Home_Streak_Non <- apply(Home_Track_Non[, c(3:9)], 1, 
                                          consec_games, value = 1)
```

### Combine Data

Combine all information we have computed into one data frame

```{r, message = FALSE}
complete <- data %>%
  left_join(Sched_Track) %>%
  left_join(Home_Track) %>%
  left_join(Home_Track_Con) %>%
  left_join(Home_Track_Non)

# Ensure everything is numeric!
complete$Games_Between_Bye <- as.numeric(complete$Games_Between_Bye)
complete$Games_Between_0 <- as.numeric(complete$Games_Between_0)
complete$Games_Between_1 <- as.numeric(complete$Games_Between_1)
complete$Games_Between_2 <- as.numeric(complete$Games_Between_2)
complete$Games_Between_3 <- as.numeric(complete$Games_Between_3)
complete$Games_Between_4 <- as.numeric(complete$Games_Between_4)
complete$Games_Between_5 <- as.numeric(complete$Games_Between_5)
complete$Games_Between_6 <- as.numeric(complete$Games_Between_6)
complete$Games_Between_7 <- as.numeric(complete$Games_Between_7)
complete$Games_Between_8 <- as.numeric(complete$Games_Between_8)
complete$Games_Between_9 <- as.numeric(complete$Games_Between_9)
complete$Games_Between_10 <- as.numeric(complete$Games_Between_10)
complete$Max_Away_Streak_Con <- as.numeric(complete$Max_Away_Streak_Con)
complete$Max_Home_Streak_Con <- as.numeric(complete$Max_Home_Streak_Con)
complete$Max_Away_Streak_Non <- as.numeric(complete$Max_Away_Streak_Non)
complete$Max_Home_Streak_Non <- as.numeric(complete$Max_Home_Streak_Non)
complete$Max_Away_Streak_Total <- as.numeric(complete$Max_Away_Streak_Total)
complete$Max_Home_Streak_Total <- as.numeric(complete$Max_Home_Streak_Total)

```

### Metrics

Calculate RMSE metrics.

```{r}
# Calculate RMSE metrics as described in the presenation.
summ <- complete %>%
  group_by(Schedule) %>%
  # Average distance accross teams for each schedule
  mutate(Average_Dist_Total = mean(Total_Dist),
         Average_Dist_NonCon = mean(NonCon_Dist),
         Average_Dist_Con = mean(Con_Dist)) %>%
  ungroup() %>%  
  group_by(Schedule, index) %>%
  mutate(
    # Distance each team is from the mean distance for the schedule squared
         Sq_Dist_Total_From_Mean = (Average_Dist_Total - Total_Dist)^2,
         Sq_Dist_NonCon_From_Mean = (Average_Dist_NonCon - NonCon_Dist)^2,
         Sq_Dist_Con_From_Mean = (Average_Dist_Con - Con_Dist)^2) %>%
  ungroup() %>%
  group_by(Schedule) %>%
  summarise(
    # We want these to be low - the square root of the mean distance from the mean accorss teams
    RMSE_Total_Dist = sqrt(mean(Sq_Dist_Total_From_Mean)),
            RMSE_NonCon_Dist =  sqrt(mean(Sq_Dist_NonCon_From_Mean)),
            RMSE_Con_Dist = sqrt(mean(Sq_Dist_Con_From_Mean)),
            
            # We want these to be high
            # The minimum distance between byes or repeat opponents across all teams for each schedule
            min_bye_dist = min(Games_Between_Bye),
                      min_0_dist = min(Games_Between_0, na.rm = TRUE),
                        min_1_dist = min(Games_Between_1, na.rm = TRUE),
                        min_2_dist = min(Games_Between_2, na.rm = TRUE),
                        min_3_dist = min(Games_Between_3, na.rm = TRUE),
                        min_4_dist = min(Games_Between_4, na.rm = TRUE),
                        min_5_dist = min(Games_Between_5, na.rm = TRUE),
                        min_6_dist = min(Games_Between_6, na.rm = TRUE),
                        min_7_dist = min(Games_Between_7, na.rm = TRUE),
                        min_8_dist = min(Games_Between_8, na.rm = TRUE),
                        min_9_dist = min(Games_Between_9, na.rm = TRUE),
                        min_10_dist = min(Games_Between_10, na.rm = TRUE),
            
            # Get the highest home and away streaks across all teams for each schedule
            max_away_streak = max(Max_Away_Streak_Total),
            max_home_streak = max(Max_Home_Streak_Total),
            max_away_streak_nc = max(Max_Away_Streak_Non),
            max_home_streak_nc = max(Max_Home_Streak_Non),
            max_away_streak_c = max(Max_Away_Streak_Con),
            max_home_streak_c = max(Max_Home_Streak_Con)) %>% ungroup()
```

```{r}
# Filter to good schedules in terms of not having byes the same opponent back to back 
final <- summ %>%
  filter(
    min_bye_dist > 0,
    min_0_dist > 0,
    min_1_dist > 0,
    min_2_dist > 0,
    min_3_dist > 0,
    min_4_dist > 0,
    min_5_dist > 0,
    min_6_dist > 0,
    min_7_dist > 0,
    min_8_dist > 0,
    min_9_dist > 0,
    min_10_dist > 0
  )
```

```{r}
# Write out the good schedules (with their schedule index number)
write.csv(final, file = "out.csv")
```
